A GPU Implementation of Capon for an Active Sonar System

### 2012-05-03 ECUA2012/gpu_capon_ecua2012.tex
la2pdf commit - 2012:05:03	15:49:59	(Thu 3.May)

The image of modern phased array active sonar systems suffer from poor lateral resolution and \gls{SNR}. This is because the lateral extent of the array is too small compared to the wavelength of the received signal, and while it is possible to trade resolution for improved \gls{SNR} by applying various weightsets to the array channels, one always end up with a compromise between the two. By introducing adaptive techniques, however, it is possible to resolve lateral details that would otherwise have been missed.

@@

Consider an $M$ element phased array. Let the sample recorded by the $m$'th channel at time instant $n$ be represented as $x_m[n]$, and the accompanying weights and delays be  $w_m[n]$ and $\Delta_m[n]$. The beamformer output $z[n]$ is then given as
\begin{align}
z[n] = \w\H[n]\X[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n-\Delta_0]\\x_1[n-\Delta_1]\\\vdots\\x_{M-1}[n-\Delta_{M-1}]}.
\end{align}
The \gls{MVDR} beamformer \cite{Capon1969} finds the set of weights that minimizes the output power, $\min E\{|z[n]|^2\}$, while ensuring unity gain in the look direction. The solution to this optimization problem is given as
\begin{gather}
\vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},
\end{gather}
where $\1$ is a row vector of ones, and $\R$ is the spatial covariance matrix. 
\begin{gather}
R[n] = \frac{1}{2N_\text{avg}+1} \sumb{n'=n-N_\text{avg}}{n+N_\text{avg}} \x[n']\x\H['n]
\end{gather}

>>>>

Modern phased array active sonar systems are often limited by its lateral resolution and \gls{SNR} of the system. This is because of the the lateral extent of the array is too small compared to the wavelength of the received signal, and while it is possible to trade resolution for improved \gls{SNR} by applying various weightsets to the array channels, one always end up with a compromise between the two. By introducing adaptive techniques, however, it is possible to resolve lateral details that would otherwise have been missed.

@@

Consider an $M$ element phased array. Let the sample recorded by the $m$'th channel at time instant $n$ be represented as $x_m[n]$, and the accompanying weights and delays be  $w_m[n]$ and $\Delta_m[n]$. The beamformer output $z[n]$ is then given as
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n-\Delta_0]\\x_1[n-\Delta_1]\\\vdots\\x_{M-1}[n-\Delta_{M-1}]}.
\end{align}
The \gls{MVDR} beamformer \cite{Capon1969} finds the set of weights that minimizes the output power, $\min E\{|z[n]|^2\}$, while ensuring unity gain in the look direction. The solution to this optimization problem is given as
\begin{gather}
\vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},
\end{gather}
where $\1$ is a row vector of ones, and $\R$ is the spatial covariance matrix. Since $\R$ is unknown, it must be estimated. First, we split the segmenting the array into subarrays where all but one element overlap, and call the output vectors
\begin{gather}
\x_l[n] = \bmat{x_l[n-\Delta_l] & x_{l+1}[n-\Delta_l] & \dots & x_{l+L-1}[n-\Delta_{l+L-1}]}\T.
\end{gather}

\begin{gather}
\tilde\R[n] = \frac{1}{M-L+1} \sumb{i=0}{M-L+1} \x_{i+k}[n]\x_{i+k}\H[n]
\end{gather}
\begin{gather}
\tilde\R[n] = \frac{1}{2N_\text{avg}+1} \sumb{n'=n-N_\text{avg}}{n+N_\text{avg}} \x[n']\x\H['n]
\end{gather}
